---
title: "Travel Safety for US citizens"
author: "project team members"
date: "1/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("knitr")
source("../source/summary-information-script.R")
source("../source/barchart.R")
```
## Summary Information
This is a summary paragraph that details important information about the data. The most recent update to this data was in `r summaryinfo$most_recent_death_date`. This data includes `r summaryinfo$num_of_locations` different cities. There are `r summaryinfo$num_of_causes` causes of death in the data with the most common cause of death being `r summaryinfo$most_common_death_cause`. The data has an average of `r summaryinfo$daily_death_avg` deaths added to the dataset each day.


source("aggregate.R")

## Aggregated Table script
**death_per_city**: By ranking the number of deaths from lowest to highest, we can observe the ranking of safety for each city. The data set for “death_per_city” will reveal the name of the city and next to it, the number of deaths that occurred in that city.

**types_of_death**: This data set reveals the Cause of Death on the first column and the occurrence of each of the causes on the second column. We would need this data to be aware of which aspect we need to be cautious of, when we travel around.


source("barchart.R")
```{r}
barchart
```

**Description** : The bar chart have list out the top 10 region that most U.S citizens die over sea. The purpose of the graph is to shows the danger of each region so people can avoid to travel to these location on the graph. As the graph shows, Mexico have approximately ten times higher than the other. 
